---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
    cache = FALSE,
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/graphics/"
)
```


# wior <img src="man/graphics/wior.svg" align="right"/>

**wior** packages several functions and utilities that make R better for handling  oceanographic data in the tidy data paradigm. It started mostly as a packaging of assorted wrappers and tricks that I wrote for my day to day work as a researcher in oceanographic sciences at the Institute of Marine Sciences, Zanzibar. 

The packages is basically developed to help marine and freshwater scientist access a large and varied format of in-situ and satellite data in easy way. In fact the package has made data access in much easy way. We are trying to remove the barrier of data access and leave scientist focus in much deeper thinking of their field rather than spending several days to understand code and syntax of language. The package contains several tools that allows scientist to get a wide array of datasets. And the funny things is that you get a tidy format result of the download, which is easy to handle in R and also to share it out of R environment. The tidy format is in form that many scientist familiar with Excel spreadsheet will find it handy.

Currently **wior** is in development but maturing. Most functions check arguments and there are some tests. However, some functions might change it's interface, and functionality can be moved to other packages, so please bear that in mind. 

## Installation

You can install development version of **wior** from github:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("lugoga/wior")
```


## Examples
In this example we easily read a CTD file, which is in `cnv` format. For those who have worked with CTD data understand the hurdle we often go through the steps to process binary data from the instrument before is in the right format for use. However, with `cnv_tibble` function from **wior** package, that process is accomplished with one line of code and you get tabular format with all the variables measured with the instruement. 
```{r}
require(wior)

ctd = wior::cnv_tibble(cnv = "e:/Data Manipulation/ctd_algoa/dstn052.cnv", interval = 10)

ctd %>% dplyr::slice(1:12)

```

You can read more in the vignettes: [Introduction to wior](https://github.com/lugoga/wior/blob/main/Intro.html).
